---
title: "linas back_ups and shit"
output: html_document
date: "2025-06-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Gaussian Fitting (nice try, but not sufficient to complex curves) -Lina
```{r}
# Load package to use a sturdy fuction for gaussian fitting, that also works with critical starting values and sloping curves

library(minpack.lm)

# Build multi gaussian sumfuction up to 3 peaks (for later use)
# note: A = amplitude (height of peak), µ = mean (peak position); sigma = standard deviation (width of peak), x = fraction
# we can exclude the normalization factor, since our data is already normalized
# for more than one peak we need to sum up the gaussian curves of multiple peaks
# for the 2nd /3rd peak we are using default values, so we can also use the function for one peak

multi_gauss <- function(x, A1, mu1, sigma1,
                          A2 = 0, mu2 = 0, sigma2 = 1,
                          A3 =0, mu3 = 0, sigma3 = 1) {
    A1 * exp(-(x - mu1)^2 / (2 * sigma1^2)) + 
    A2 * exp (-(x - mu2)^2 / (2* sigma2^2)) + 
    A3 * exp(-(x - mu3)^2 / (2* sigma3^2))
  }

# build fuction to determin of number of peaks on raw data (for later use)
# threshold t is the minimal peak hight we are acepting (3% of maximal intensity, the rest ist noise) <- has proven to be a good limit
# note: diff() computes the slope between to points, sign() then gives the sign of the slope +1/0/-1, if slope changes from positive to negative diff() between +1 and -1 = -2

t = max(y) * 0.03 
find_peaks <- function(y, threshold = t) { 
  peaks <- which(diff(sign(diff(y))) == -2) + 1 #(+1 shifts index back)
  peaks[y[peaks]>= threshold] # [peaks] = peak intensität <- only those over thresohold
}

# create list for results

fit_results <- list()

# x-achsis for fited curves (fraction 1-25)

x <- 1:25

# fitting of every protein (Ctrl and RNAse) using for-loops
# for loop that goes through every protein/row, extracts the Ctrl values and RNAse and saves them in seperate vectors

for (p in seq_len(nrow(MS_Table_Norm))) {
  row_vals <- as.numeric(MS_Table_Norm [p,])
  ctrl_vals <- row_vals[seq(1, 50, by =2)]
  rnase_vals <- row_vals[seq(2, 50, by=2)]
  
  #create emty list for results of this protein
  
  results_row <- list()
  
  # second for loop that first works with Ctrl, then RNAse values of protein
  # with cond and vector c("Ctrl", "RNAse"), we are computing the for loop to run exactly two times, first under condition "Ctrl" then under condition "RNAse"
  # in the first loop under "Ctrl" the ctrl_values are used, then in the second loop the rnase_vals
  #
  for (cond in c("Ctrl", "RNAse")) {
    y <- if (cond == "Ctrl") ctrl_vals else rnase_vals 
    
    # finding peaks, if no peaks are detected --> fallback mechanism: find maximum intensity and asign fraction as peak position
    
    peaks <- find_peaks(y)
    if(length(peaks) == 0) {
      peaks <- which.max(y)
    }
    
    #determine how many peaks have been found, but use 3 at a max
    
    n_peaks <- min(length(peaks),3)
    
    # extract intensity values for each peak as starting values for A (y[peaks]) and fraction of each peak as starting value for mu (x[peaks])
    
    y_peaks <- y[peaks]
    x_peaks <- x[peaks]
    
    #define start values for up to three peaks using if clauses; sigma = 2 was an estimate that has proven to be a good starting value, since there are 25 fractions
    
    starts <- list(A1 = y_peaks[1], mu1 = x_peaks[1], sigma1 = 2)
    if (n_peaks >= 2) {
      starts <- c(starts, list(A2 = y_peaks[2], mu2 = x_peaks[2], sigma2 = 2)) }
      else {
        starts <- c(starts, list(A2 = 0, mu2 = 0, sigma2 = 2))}
    if (n_peaks == 3) {
      starts <- c(starts, list(A3 = y_peaks[3], mu3 = x_peaks[3], sigma3 = 2)) }
      else {
        starts <- c(starts, list(A3 = 0, mu3 = 0, sigma3 = 2))}

  
    # fitting with our multi gaussian fuction and try(), collect fit data in vector "fit"
    # using if clauses to go through every possibility: 1, 2,  or 3 peaks, and build a model with 1 up to three gaussian curves
    # we can't use multi_gauss with one peak, so we fit those with a normal gauss function (has proven to be necessary)
    # nlsLM = function that fits our gaussian curves formular /function we are giving "y ~..." (this is a nonlinear fitting fuction)
    # we need a lower and a upper limit, since in extreme cases, the estimation can be unrealistic (proven)
    # we are using "maxiter = 500" , meaning that the function will try to fit until 500 iterations and not break of earlier/longer search time (proven to be needed, since the function would terminate many estimations, that could be successful)
    # not all fits ar going to be successful:
    # "warnOnly" guarantees that the code keeps running, even if the fitting of a protein is not successful (happens for a few proteins)
    # function try() is another important protective measure, to ensure, that if "try" was not successful, the code keeps running, and there is only an error message for this protein (silent = TRUE <- error message not shown in console) 
    
    if (n_peaks == 1) {
      fit <- try(nlsLM(y ~ A1 * exp(-((x - mu1)^2) / (2 * sigma1^2)),
                       start = list(A1 = y_peaks[1], mu1 = x_peaks[1], sigma1 = 2),
                                    lower = c(A1 = 0, mu1 = 1, sigma1 = 0.3),
                                    upper = c(A1 = Inf, mu1 = 25, sigma1 = 5),
                       control = list(maxiter = 500, warnOnly = TRUE)
                       ), 
                 silent = TRUE)
      } 
    else if (n_peaks == 2) {
      fit <- try(nlsLM(y ~ multi_gauss(x, A1, mu1, sigma1, A2, mu2, sigma2),
                       start = list(
                         A1 = y_peaks[1], mu1 = x_peaks[1], sigma1 = 2,
                         A2 = y_peaks[2], mu2 = x_peaks[2], sigma2 = 2),
                       lower = c(A1 = 0, mu1 = 1, sigma1 = 0.3,
                                 A2 = 0, mu2 = 1, sigma2 = 0.3),
                       upper = c(A1 = Inf, mu1 = 25, sigma1 = 5,
                                 A2 = Inf, mu2 = 25, sigma2 = 5),
                       control = list(maxiter = 500, warnOnly = TRUE)
                       ), 
                 silent = TRUE)
      } 
    else {fit <- try(nlsLM(y ~ multi_gauss(x, A1, mu1, sigma1, A2, mu2, sigma2, A3, mu3, sigma3),
                           start = list(
                              A1 = y_peaks[1], mu1 = x_peaks[1], sigma1 = 2,
                             A2 = y_peaks[2], mu2 = x_peaks[2], sigma2 = 2,
                             A3 = y_peaks[3], mu3 = x_peaks[3], sigma3 = 2),
                           lower = c(A1 = 0, mu1 = 1, sigma1 = 0.3,
                                     A2 = 0, mu2 = 1, sigma2 = 0.3,
                                     A3 = 0, mu3 = 1, sigma3 = 0.3),
                           upper = c(A1 = Inf, mu1 = 25, sigma1 = 5,
                                     A2 = Inf, mu2 = 25, sigma2 = 5,
                                     A3 = Inf, mu3 = 25, sigma3 = 5),
                           control = list(maxiter = 500, warnOnly = TRUE)
                           ), 
                     silent = TRUE)
   }
    
    # collect results for this protein in results_row and add to fit_results list
    # if clauses for when there is no "try-error" message (successful fit) for protein and when there is an error (n)
    # "coef" extracts parameter values (A1, mu1, ../ coefficients)
    # give parameters name in loist results_row and assign computed values 
    # "paste0" takes a text frgment given and puts them together without spaces
    # [[]] needed to access a specific value, not a list
    # assign new variable "Fit_OK_ " to result list to show successful and not successfully fitted proteins
    
    if(!inherits(fit, "try-error")) {
      coef_vals <- coef(fit)
      for (i in names(coef_vals)) {
        results_row[[paste0(i,"_", cond)]] <- coef_vals[i]
      }
      results_row[[paste0("Fit_OK_", cond)]] <- TRUE
      results_row[[paste0("n_peaks_", cond)]] <- n_peaks
    }
    else {
      results_row[[paste0("Fit_OK_", cond)]] <- FALSE
      results_row[[paste0("n_peaks_", cond)]] <- 0
    }
  }
  # close for loop, and save protein data in list fit_results
  fit_results[[p]] <- results_row
}

# make data.frame out of list (quite complicated, cause not every protein has the same amount of values)
# going through data for each protein x in fit_results
# if there is no value for a specific field/parameter (A1, mu1,...) we are assigning "NA" (%in% = "is included in")

rows_df <- lapply(fit_results, function(x){
  fields <- c("A1", "mu1", "sigma1", "A2", "mu2", "sigma2", "A3", "mu3", "sigma3")
  for (cond in c("Ctrl", "RNAse")) {
    for (f in fields) {
      key <- paste0(f, "_", cond)
      if(!(key %in% names(x))) {
        x[[key]] <- NA
      }
    }
  }
  as.data.frame(x)
})

#assemble all data in one dataframe
# problem: rows_df is a list of one row dataframes (since we used lapply), we need to split up the list with do.call, and bind all dataframes with rbind
MS_Data_Fitted <- do.call(rbind, rows_df)
rownames(MS_Data_Fitted) <- rownames(MS_Table_Norm)

Protein_IDs_fit_error <- rownames(subset(MS_Data_Fitted, Fit_OK_Ctrl == FALSE | Fit_OK_RNAse == FALSE))
```

### Plotting Protein with both normalised and gaussian fit curves - Lina
```{r}
plot_protein_inkl_gauss <- function(x) {
  protein_of_interest <- x 
  protein_norm_data <- MS_Table_Norm[rownames(MS_Table_Norm) == protein_of_interest, ]
  ctrl_values_norm_data <- as.numeric(protein_norm_data[seq(1, 49, by=2)])
  rnase_values_norm_data <- as.numeric(protein_norm_data[seq(2, 50, by=2)])
  fractions <- 1:25
  
  if (!(protein_of_interest %in% Protein_IDs_fit_error)) {
    protein_gauss_data <- MS_Data_Fitted[rownames(MS_Data_Fitted) == protein_of_interest, , drop = FALSE]
    x_vals <- seq(1, 25, by = 0.1)
    
    gauss_curve <- function(protein_gauss_data, condition, x) {
       y <- rep(0, length(x))
       n_peaks <- protein_gauss_data[[paste0("n_peaks_", condition)]]
       for (i in 1:n_peaks) {
         A <- protein_gauss_data[[paste0("A", i, "_", condition)]]
         mu <- protein_gauss_data[[paste0("mu", i, "_", condition)]]
         sigma <- protein_gauss_data[[paste0("sigma", i, "_", condition)]]
         y <- y + A * exp(-((x - mu)^2) / (2 * sigma^2))
         }
       return(y)
    }
    y_ctrl <- gauss_curve(protein_gauss_data, "Ctrl", x_vals)
    y_rnase <- gauss_curve(protein_gauss_data, "RNAse", x_vals)
    } 
  else {
    x_vals <- NULL
    y_ctrl <- y_rnase <- NULL
    }

  plot(fractions, ctrl_values, type="o", pch=16, lwd=1, lty = 1, col="green", ylim=range(c(ctrl_values, rnase_values), na.rm = TRUE),
       xlab="Fraction", ylab="Intensity", main= protein_of_interest)
  lines(fractions, rnase_values, type="o",pch =16, lwd=1, lty = 1,  col="red")
  
  if (!is.null(x_vals)) {
    lines(x_vals, y_ctrl, col="green", lwd=2)
    lines(x_vals, y_rnase, col="red", lwd=2)
  }
  
  legend("topright", 
         legend=c("Ctrl (raw)", "RNase (raw)", "Ctrl (fit)", "RNase (fit)"),
         col=c("green", "red", "green", "red"), 
         lty=c(2,2,1,1), pch=c(16,16,NA,NA), lwd=c(1,1,2,2))
}

plot_protein_inkl_gauss("ANKUB_HUMAN") #fill in protein that you wanna see
```

## First Idea fpr Shift Calculation - Using Means of Peak Positions (Heights not included!) - Lina 
```{r}
# Berechnung Average Peak Shift 
  means_peak_position_Ctrl <- numeric(nrow(Peak_Data))
  means_peak_position_RNAse <- numeric(nrow(Peak_Data))
  shift_length <- numeric(nrow(Peak_Data))
  
for (p in seq_len(nrow(Peak_Data))) {
  
  positions_Ctrl <- as.numeric(sapply(1:6, function(i) Peak_Data[p, paste0("peak", i, "_position_Ctrl")]))
  means_peak_position_Ctrl[p] <- mean(positions_Ctrl, na.rm = TRUE)
  
  positions_RNAse <- as.numeric(sapply(1:6, function(i) Peak_Data[p, paste0("peak", i, "_position_RNAse")]))
  means_peak_position_RNAse[p] <- mean(positions_RNAse, na.rm = TRUE)
  
  shift_length[p] <- means_peak_position_Ctrl[p] - means_peak_position_RNAse[p]
}

  Shift_Data <- data.frame(
  means_peak_position_Ctrl = means_peak_position_Ctrl,
  means_peak_position_RNAse = means_peak_position_RNAse,
  shift_length = shift_length,
  row.names = rownames(Peak_Data))

head(Shift_Data)
```


### FDR Correction of T-Text - Lina (decreased our hit-rate to much)
```{r}
# create empty columns for statistical test results 
Shift_Data$shapiro_p <- NA
Shift_Data$normal_distributed <- NA
Shift_Data$p_value_ttest_filtered <- NA
Shift_Data$significant_ttest_filtered <- NA

# iterate over all proteins 
for (protein in rownames(Shift_Data_for_t)) {
  
  # extract shift distances for each replicate
  shift_values <- as.numeric(Shift_Data_for_t[protein, c("shift_distance_rep1", "shift_distance_rep2", "shift_distance_rep3")])
  
  # make sure that no NAs are present
  if (all(!is.na(shift_values))) {
    
    # check if all values are identical (no variance)
    if (length(unique(shift_values)) == 1) {
      
      # set normal distribution to TRUE if no variance is present
      Shift_Data[protein, "shapiro_p"] <- NA
      Shift_Data[protein, "normal_distributed"] <- TRUE
      
    } else {
      
      # perform Shapiro-Wilk test to check for normal distribution
      shapiro_result <- shapiro.test(shift_values)
      Shift_Data[protein, "shapiro_p"] <- shapiro_result$p.value
      Shift_Data[protein, "normal_distributed"] <- shapiro_result$p.value >= 0.05
    }
    
    # perfrom one-sided t-test if data is normal distributed
    if (Shift_Data[protein, "normal_distributed"] == TRUE) {
      
      test_result <- t.test(shift_values, mu = 1, alternative = "greater")
      
      Shift_Data[protein, "p_value_ttest_filtered"] <- test_result$p.value
      Shift_Data[protein, "significant_ttest_filtered"] <- ifelse(test_result$p.value < 0.05, TRUE, FALSE)
      
    } else {
      # no t-test applied if data is not normal distributed
      Shift_Data[protein, "p_value_ttest_filtered"] <- NA
      Shift_Data[protein, "significant_ttest_filtered"] <- NA
    }
  }
}

# summary of proteins regarding normal distribution
table(Shift_Data$normal_distributed, useNA = "ifany")

# summary of proteins regarding significant of t-test
table(Shift_Data$significant_ttest_filtered, useNA = "ifany")

# Jetzt die Multiple Testing Correction (FDR)

# Neue Spalten für FDR-Korrektur anlegen
Shift_Data$p_value_ttest_fdr <- NA
Shift_Data$significant_ttest_fdr <- NA

# Nur p-Werte ohne NAs extrahieren
pvals <- Shift_Data$p_value_ttest_filtered
pvals_adj <- p.adjust(pvals, method = "BH")

# Ergebnisse zurückschreiben
Shift_Data$p_value_ttest_fdr <- pvals_adj
Shift_Data$significant_ttest_fdr <- ifelse(pvals_adj < 0.05, TRUE, FALSE)

# Übersicht FDR-Korrektur
table(Shift_Data$significant_ttest_fdr, useNA = "ifany")

head(Shift_Data)
```

#Linear Regression with CoM instead of max-peak-position
```{r}
# Correlation
cor.test(Protein_Data_for_LR$CoM_RNase,
         Protein_Data_for_LR$MolecularWeight_kDa,
         method = "spearman", use = "complete.obs")

# Linear Regression
lm_model_CoM <- lm(MolecularWeight_kDa ~ CoM_RNase, data = Protein_Data_for_LR)
summary(lm_model_CoM)

#Scatter Plot
ggplot(Protein_Data_for_LR %>% filter(MolecularWeight_kDa <= 1000), aes(x = CoM_RNase, y = MolecularWeight_kDa)) +
  geom_point(aes(color = "All analysed Proteins")) +
  geom_point(data = CORUM_comlexes,aes(x = CoM_RNase, y = MolecularWeight_kDa, color = "Proteins in Complex (CORUM)"),alpha = 0.5)  +
  geom_abline(intercept = coef(linear_reference)[1], 
              slope = coef(linear_reference)[2], 
              color = "darkred", size = 1) +
  geom_point(data = reference_proteins, aes(x = x, y = y), color = "darkred", size = 3) +
  geom_text(data = reference_proteins[c(1), ], aes(x = x, y = y, label = name),
            vjust = -1,   hjust = 0.6, color = "darkred", size = 4, fontface = "bold") +
  geom_text(data = reference_proteins[c(5), ], aes(x = x, y = y, label = name),
            vjust = 0.3 ,   hjust = 1.2, color = "darkred", size = 4, fontface = "bold") +
  geom_text(data = reference_proteins[c(2, 3, 4), ], aes(x = x, y = y, label = name),
            vjust = 0.3,   hjust = -0.2, color = "darkred", size = 4, fontface = "bold") +
  theme_classic() + theme(legend.position =  c(0.2, 0.85)) +
  scale_color_manual(values = c("All analysed Proteins" = "gray75", "Proteins in Complex (CORUM)" = "lightcoral")) +
  labs(x = "CoM_RNase", y = "Molecular Weight (kDa)",  color = "Protein Category", title = "Molecular Weight vs. CoM_RNase")
```


## Cluster Try - Cihan
```{r}
# 1) Filter only the RBPs relevant in Mitosis 
filtered_shift_Mitosis <- Shift_Data[rownames(Shift_Data) %in% RBPs_Mitosis, ]
filtered_peak_Mitosis <- Peak_Data[rownames(Peak_Data) %in% RBPs_Mitosis, ]

# 2) Selecting only the interesting Ctrl peak columns 
ctrl_peak_cols_cluster <- grep("_Ctrl$", colnames(filtered_peak_Mitosis), value = TRUE)
merged_filterd_mitosis <- filtered_peak_Mitosis[, ctrl_peak_cols_cluster]

# 3) Clean data
merged_filterd_mitosis[is.na(merged_filterd_mitosis)] <- 0

clean2_merged_filterd_mitosis <- merged_filterd_mitosis[, apply(merged_filterd_mitosis, 2, function(col) sd(col) != 0)]

# 4) Prepare matrix for clustering
clean2_merged_filterd_mitosis <- as.matrix(clean2_merged_filterd_mitosis)

# 5) Distance and hierarchical clustering
dist_matrix <- dist(clean2_merged_filterd_mitosis)
hc_ward <- hclust(dist_matrix, method = "ward.D2")

# 6) Plot dendrogram
par(mfrow = c(1,1), mar = c(5, 4, 4, 2))
plot(hc_ward,
     labels = FALSE,
     main = "Ward's Method",
     xlab = "", ylab = "Height",
     cex = 1.2,
     hang = -1)

# 7) Assign clusters (e.g., k = 6)
cluster_membership <- cutree(hc_ward, k = 10)

# 8) Store clustering result in a data frame
protein_clusters <- data.frame(
  Protein = rownames(clean2_merged_filterd_mitosis),
  Cluster = cluster_membership
)

# 9) Sort by cluster number
protein_clusters_sorted <- protein_clusters[order(protein_clusters$Cluster), ]

table(protein_clusters_sorted$Cluster)

# 11) Optional: Plot proteins in a selected cluster, e.g., cluster 5
cluster5_proteins <- protein_clusters_sorted[protein_clusters_sorted$Cluster == 8, "Protein"]

for (r in cluster5_proteins) {
  plot_protein(r)
}







library(cluster)

# Distanzmatrix und hclust musst du vorher schon haben:
# dist_matrix <- dist(peak_matrix)
# hc_ward <- hclust(dist_matrix, method = "ward.D2")

sil_scores <- numeric()

# Schleife für k = 2 bis 10
for (k in 2:10) {
  cluster_assignments <- cutree(hc_ward, k = k)
  sil <- silhouette(cluster_assignments, dist_matrix)
  sil_scores[k] <- mean(sil[, 3])  # Durchschnittliche Silhouettenbreite
}

# Plot
plot(2:10, sil_scores[2:10], type = "b", pch = 19,
     xlab = "Anzahl Cluster (k)",
     ylab = "Durchschnittliche Silhouettenbreite",
     main = "Silhouettenanalyse für hierarchisches Clustering")



```

## Second Chunk of Cluster - Cihan
```{r}
# 1. peak1_position_Ctrl & peak1_height_Ctrl vorbereiten
peak1_df <- data.frame(
  Protein = rownames(Peak_Data),
  peak1_position_Ctrl = Peak_Data$peak1_position_Ctrl,
  peak1_height_Ctrl = Peak_Data$peak1_height_Ctrl
)

# 2. Merge mit den Clusterzuordnungen
peak_data_with_clusters <- merge(
  protein_clusters_sorted,
  peak1_df,
  by = "Protein"
)

# 3. Leere Liste zur Speicherung der gefilterten Cluster initialisieren
filtered_clusters <- list()

# 4. Schleife über alle Cluster
for (cl in unique(peak_data_with_clusters$Cluster)) {
  cluster_subset <- subset(peak_data_with_clusters, Cluster == cl)
  
  # Rundung auf ganze Zahlen zur Peakgruppierung
  peak_pos <- round(cluster_subset$peak1_position_Ctrl)
  
  # Häufigste Peakposition
  freq_table <- table(peak_pos)
  if (length(freq_table) == 0) next  # überspringe leeren Cluster
  top_peak <- as.numeric(names(freq_table)[which.max(freq_table)])
  
  # Erlaubte Positionen (±1)
  allowed_range <- (top_peak - 1):(top_peak + 1)
  
  # Filter: Position innerhalb Range & Höhe mindestens 5
  keep <- peak_pos %in% allowed_range & cluster_subset$peak1_height_Ctrl >= 5
  filtered <- cluster_subset[keep, ]
  
  # Speichern
  filtered_clusters[[paste0("Cluster_", cl)]] <- filtered
}

# 5. Alle gefilterten Cluster zusammenführen
final_filtered_proteins <- do.call(rbind, filtered_clusters)

# 6. Export als CSV
write.csv(final_filtered_proteins, "RBP_FilteredByPeakPosition.csv", row.names = FALSE)

# 7. Übersicht anzeigen
table(final_filtered_proteins$Cluster)
#

## Plotten, nie streichen
cluster3_proteins_final <- final_filtered_proteins[final_filtered_proteins$Cluster == 8, "Protein"]

for (r in cluster3_proteins_final) {
  plot_protein(r)
}




```


## doing a PCa analysis in order to look at the distribution of RBPS => could be applyed to selected RBPs from mytosis - Sofia
```{r}
# Rows = proteins, Columns = the 26 features you mentioned
# First, remove rows with any NA values
Peak_Data_clean <- na.omit(Peak_Data)

# Remove columns with zero variance
Peak_Data_clean_clean <- Peak_Data_clean[, apply(Peak_Data_clean, 2, function(col) sd(col, na.rm = TRUE) != 0)]

# Perform PCA (center and scale the data)
pca_result <- prcomp(Peak_Data_clean_clean, center = TRUE, scale. = TRUE)

# Extract the scores for the first two principal components
pc_scores <- as.data.frame(pca_result$x[, 1:2])

# Add protein names as a column (assuming they are rownames)
pc_scores$Protein <- rownames(pc_scores)

# Create a new column that marks whether the protein is in t_test_positives (taken from further steps of the analysis)

pc_scores$Significant <- ifelse(pc_scores$Protein %in% t_test_positives, "Yes", "No")

# Plot the PCA
ggplot(pc_scores, aes(x = PC1, y = PC2, color = Significant)) +
  geom_point(alpha = 0.7, size = 1.5) +
  scale_color_manual(values = c("Yes" = "red", "No" = "black")) +
  theme_minimal() +
  labs(title = "PCA of Protein Features (highlighting significant proteins)",
       x = "PC1", y = "PC2", color = "Significant")


```


# completly new just trying to see if the normalisation was the reason our PC analysis wasnt working - Sofia 

```{r}
                        # neue Normalization that maybe works better: 


# Log2 transformation (adding 1 to avoid log(0))
MS_log <- log2(MS_Table_Averages + 1)

# Row-wise z-score scaling: center and scale each protein across all samples (fractions)
MS_scaled <- t(scale(t(MS_log)))

rownames(MS_scaled) <- rownames(MS_Table_Averages)

# Optional: check a few rows
head(round(MS_scaled, 2))

#filter proteins with low variance 
# 1. Compute variance across proteins (handle NAs)
var_filter <- apply(MS_scaled, 1, var, na.rm = TRUE)

# 2. Compute the 75th percentile of variance, handling NAs in var_filter
threshold <- quantile(var_filter, 0.75, na.rm = TRUE)

# 3. Filter for high-variance proteins
MS_highvar <- MS_scaled[var_filter > threshold, ]

# If any are missing, fix them by assigning unique names

rownames(MS_highvar) <- make.names(rownames(MS_highvar), unique = TRUE)



                         # Linas code zu Peak characterization just applyed to my code 



# fractions = x
x <- 1:25

# define treshold for peak and build function for peak detection
find_peaks <- function(y, threshold = max(y) * 0.03) { 
  peaks <- which(diff(sign(diff(y))) == -2) + 1 #(+1 shifts index back)
  peaks[y[peaks]>= threshold] 
}

# create empty list for results
peak_results <- list()

# for loop over every protein
for (p in seq_len(nrow(MS_highvar))) {
  
  # extract values for Ctrl and RNase
  row_vals <- as.numeric(MS_highvar [p,])
  ctrl_vals <- row_vals[seq(1, 50, by =2)]
  rnase_vals <- row_vals[seq(2, 50, by=2)]
  
  # create empty list for results
  results_row <- list()
  
  # for loop working with Ctrl values first, then with RNase values
  for (cond in c("Ctrl", "RNase")) {
    y <- if (cond == "Ctrl") ctrl_vals else rnase_vals 
    
    # find peaks with earlier build function, if no peak is detected peak1 is max intensity
    peaks <- find_peaks(y)
    if(length(peaks) == 0) {
      peaks <- which.max(y)
    }
    
    # store numer of peaks, but with a maximum of 6 
    n_peaks <- min(length(peaks), 6)
    
    # extract peak height y_peaks and peak position x_peaks
    y_peaks <- y[peaks]
    x_peaks <- x[peaks]

    # store different results in list 
    results_row[[paste0("n_peaks_", cond)]] <- n_peaks
    
    for(n in seq_len(n_peaks)) {
    results_row[[paste0("peak", n, "_height_", cond)]] <- y_peaks[n]
    results_row[[paste0("peak", n, "_position_", cond)]] <- x_peaks[n]
    }
  }
  # store protein results in lisr
  peak_results[[p]] <- results_row
}

# befor we can store data in new dataframe, we need to assign NA to all columns that have no values 
rows_df <- lapply(peak_results, function(x){
  columns <- c("peak1_height", "peak1_position", "peak2_height", "peak2_position", "peak3_height", "peak3_position", "peak4_height", "peak4_position", "peak5_height", "peak5_position", "peak6_height", "peak6_position")
  for (cond in c("Ctrl", "RNase")) {
    for (c in columns) {
      key <- paste0(c, "_", cond)
      if(!(key %in% names(x))) {
        x[[key]] <- NA
      }
    }
  }
  as.data.frame(x)
})

# store results in new DataFrame (note: rows_df is a list of dataframes so we need "do.call(r.bind,...)")
Peak_Data2 <- do.call(rbind, rows_df)
rownames(Peak_Data2) <- rownames(MS_highvar)

# reorder columns of Peak_Data2
order <- c("n_peaks_Ctrl", "n_peaks_RNase")

for (i in 1:6) {
  order <- c(order, 
                 paste0("peak", i, "_height_Ctrl"),
                 paste0("peak", i, "_position_Ctrl"),
                 paste0("peak", i, "_height_RNase"),
                 paste0("peak", i, "_position_RNase"))
}

Peak_Data2 <- Peak_Data2[, order]

# show head of PeakData to check 
head(Peak_Data2)

colnames(Peak_Data2)




                          # PCa jetzt auf neue daten runnen + vergleich zu RBPs 


# Load required package
library(ggplot2)


# Assume your data is in a data frame called 'df'
# Rows = proteins, Columns = the 26 features you mentioned
# First, remove rows with any NA values
Peak_Data_clean <- na.omit(Peak_Data2)

# Remove columns with zero variance
Peak_Data_clean_clean <- Peak_Data_clean[, apply(Peak_Data_clean, 2, function(col) sd(col, na.rm = TRUE) != 0)]

# Perform PCA (center and scale the data)
pca_result <- prcomp(Peak_Data_clean_clean, center = TRUE, scale. = TRUE)

# Extract the scores for the first two principal components
pc_scores <- as.data.frame(pca_result$x[, 1:2])

# Add protein names as a column (assuming they are rownames)
pc_scores$Protein <- rownames(pc_scores)

# Create a new column that marks whether the protein is in t_test_positives (taken from further steps of the analysis)

pc_scores$Significant <- ifelse(pc_scores$Protein %in% t_test_positives, "RBPs", "No")

# Plot the PCA
ggplot(pc_scores, aes(x = PC1, y = PC2, color = Significant)) +
  geom_point(alpha = 0.7, size = 1.5) +
  scale_color_manual(values = c("RBPs" = "red", "No" = "black")) +
  theme_minimal() +
  labs(title = "PCA of Protein Features (highlighting significant proteins)",
       x = "PC1", y = "PC2", color = "Significant")


```




#  kmeans clustering nur auf RBPs in mitosis ( no selection of spefic parameters) - Sofia 
```{r}
# Filter only the RBPs relevant in Mitosis 
filtered_data_shift <- Shift_Data[rownames(Shift_Data) %in% RBPs_Mitosis, ]

# Load your dataset (CSV or kmeans frame)
kmeans <- filtered_data_shift # Replace with your file or data frame


# Select 4 specific columns by name or position


K <- kmeans[, c("CoM_Ctrl", "CoM_RNase", "shift_distance")]  # Replace with column names

K <- na.omit(K)          # Remove rows with any NA
K <- as.matrix(K)        # Ensure it’s still a numeric matrix


# Convert to matrix
K <- as.matrix(K)

# K-means clustering
k <- 3                    #number of clusters 
set.seed(42)
kmeans_result <- kmeans(K, centers = k) # here i put in my data 

# Reduce to 2D with PCA
pca_results <- prcomp(K, scale. = TRUE)
K_2D <- pca_results$x[, 1:2]

# Plot clusters
plot(K_2D, col = kmeans_result$cluster, pch = 19,
     main = "K-means Clustering (PCA-reduced 2D)",
     xlab = "PC1", ylab = "PC2")
legend("topright", legend = paste("Cluster", 1:k), col = 1:k, pch = 19)





```
